---
title: "Spam nnet"
author: "Spencer Braun"
date: "6/5/2020"
output:
  html_document: default
  number_sections: no
  highlight: haddock
  theme: journal
  toc: yes
  toc_depth: 3
  toc_float: yes
---
---

```{r, message=FALSE}
if (!require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse) 
pacman::p_load(here) 
pacman::p_load(glue) 
pacman::p_load(class) 
pacman::p_load(nnet) 
pacman::p_load(ROCR) 
pacman::p_load(hrbrthemes)

theme_set(hrbrthemes::theme_ipsum())
```


```{r}
spam.train <- readr::read_csv(here::here("data", "spam_train.csv"))
spam.test <- readr::read_csv(here::here("data", "spam_test.csv"))

columns <- c('word_freq_make',
'word_freq_address',
'word_freq_all',
'word_freq_3d',
'word_freq_our',
'word_freq_over',
'word_freq_remove',
'word_freq_internet',
'word_freq_order',
'word_freq_mail',
'word_freq_receive',
'word_freq_will',
'word_freq_people',
'word_freq_report',
'word_freq_addresses',
'word_freq_free',
'word_freq_business',
'word_freq_email',
'word_freq_you',
'word_freq_credit',
'word_freq_your',
'word_freq_font',
'word_freq_000',
'word_freq_money',
'word_freq_hp',
'word_freq_hpl',
'word_freq_george',
'word_freq_650',
'word_freq_lab',
'word_freq_labs',
'word_freq_telnet',
'word_freq_857',
'word_freq_data',
'word_freq_415',
'word_freq_85',
'word_freq_technology',
'word_freq_1999',
'word_freq_parts',
'word_freq_pm',
'word_freq_direct',
'word_freq_cs',
'word_freq_meeting',
'word_freq_original',
'word_freq_project',
'word_freq_re',
'word_freq_edu',
'word_freq_table',
'word_freq_conference',
'char_freq_semi_col',
'char_freq_paren',
'char_freq_bracket',
'char_freq_excl_pt',
'char_freq_dlr_sign',
'char_freq_pound',
'capital_run_length_average',
'capital_run_length_longest',
'capital_run_length_total',
'spam_label')

names(spam.train) <- columns
names(spam.test) <- columns

set.seed(123)
rows <- sample(nrow(spam.train))
spam.train <- spam.train[rows, ]

train.std <- scale(spam.train)
test.std <- scale(spam.test)

```


Test number of units for a single layer

```{r}
set.seed(123)

unit_test <- matrix(NA, 10, 2)
for (units in 1:10) {
  spam.net <- nnet(spam_label ~ ., data=spam.train, size = units, rang = 0.5, 
                 decay = 0.01, maxit = 400, trace=FALSE)
  pred.spam <- predict(spam.net, spam.test, type = "raw")
  class.spam <- ifelse(pred.spam > 0.5, 1, 0)
  unit_test[units,1] <- mean((spam.test$spam_label - pred.spam)^2)
  unit_test[units,2] <- mean(spam.test$spam_label != class.spam)
}

unit_test
best_unit <- which.min(unit_test[,2])
```

The model with `r best_unit` hidden units achieves the lowest test misclassification error rate of `r round(unit_test[best_unit,2],4)*100`%. 


Test different values of weight decay

```{r}
set.seed(123)

# Wider grid search
# decay_values = sapply(seq(0,5, by=0.5), function(x) 10^(-1*x))

# More granular grid search around best found in prior round
decay_values = sapply(seq(0,3, length=10), function(x) 10^(-1*x))
search_matrix <- matrix(NA, 10, length(decay_values))

run.net <- function(decay, maxit, units) {
  spam.net <- nnet(spam_label ~ ., data=spam.train, size = units, rang = 0.5, 
                 decay = decay, maxit = maxit, trace=FALSE)
  pred.spam <- predict(spam.net, spam.test, type = "raw")
  class.spam <- ifelse(pred.spam > 0.5, 1, 0)
  mean(spam.test$spam_label != class.spam)
}

#pb <- txtProgressBar(min = 1, max = length(decay_values), style = 3)

for (i in 1:length(decay_values)) {
  for (iter in 1:10) {
    search_matrix[iter, i] <- run.net(decay_values[i], 400, best_unit)
  }
  #setTxtProgressBar(pb, i)
}

best_decay <- decay_values[which.min(apply(search_matrix, 2, mean))]
#wider grid search best 10^(-1.5) #0.03989570
```

The final model obtained from the tuning process has `r best_unit` units and a weight decay parameter of `r best_decay`. The estimated misclassification error achieved with this model is `r round(min(apply(search_matrix, 2, mean)),4)*100`%.

Require FP rate less than 1%

```{r}
set.seed(123)

spam.net <- nnet(spam_label ~ ., data=spam.train, size = 8, rang = 0.5, 
                 decay = best_decay, maxit = 400, trace=FALSE)
pred.spam <- predict(spam.net, spam.test, type = "raw")


pred1 <- prediction(pred.spam, spam.test$spam_label)
perf1 <- performance(pred1,"tpr","fpr")
plot(perf1, colorize=FALSE, main="ROC for Spam Test")
roc_df <- data.frame(cut=perf1@alpha.values[[1]], 
                      fpr=perf1@x.values[[1]], 
                      tpr=perf1@y.values[[1]])
roc_df <- roc_df %>% 
  arrange(desc(tpr)) %>%
  filter(fpr < 0.01)

thresh.class <- ifelse(pred.spam >= roc_df[1,'cut'], 1, 0)

thresh.table <- table(thresh.class, spam.test$spam_label)
thresh.table
```

With the threshold chosen from the ROC curve, we get a total error rate of `r round(mean(thresh.class != spam.test$spam_label),4)*100`% and a false positive rate (misclassified good emails) of `r round(thresh.table[2,1] / sum(thresh.table[,1]),4)*100`%.