---
title: "Varying Coefficient Model"
author: "Spencer Braun"
date: "6/5/2020"
output: 
  html_document:
    highlight: haddock
    theme: journal
    number_sections: no
    toc: no
---


```{r, message=FALSE}
library(tidyverse)
library(reshape2)
library(ggplot2)
library(e1071)
library(MASS)
library(mvtnorm)
library(glmnet)
library(splines)
library(gam)
library(here)
library(hrbrthemes)

theme_set(hrbrthemes::theme_ipsum())
```

* Build a natural cubic spline basis for the time data. 
* Coefficients of $x_1,x_2$ as a function of the splines, $\beta=\mathbf{H} \theta$ for the basis matrix H of natural cubic splines 
* Fit a regression on the filtered data $x^* = H^Tx$, which gives us the value of $\hat{\theta}$, the regression coefficients on the filtered x values 
* Calculate the smooth coefficient curves from $\hat{\beta}(t)=h(t)^T \hat{\theta}$. Since the H matrix is a basis with respect to t, we end up with coefficients that vary as a function of time.

```{r, message=FALSE, warning=FALSE}

vcdata <- read_csv(here::here("data", "vcdata.csv"))
vcdata <- vcdata %>% dplyr::select(- X1)

## Create spline basis
df = 10
N <- ns(vcdata$t, df = 10, intercept = TRUE)

## Create theta esimates
X <- as.data.frame(cbind(rep(1, nrow(vcdata)), vcdata$x.1, vcdata$x.2))
colnames(X) <- c("int", "X1", "X2")
fit <- lm(vcdata$y ~ N:(int + X1 + X2) - 1, data=X)
coef <- fit$coefficients

# plot(vcdata$t, vcdata$y, type='l')
# lines(vcdata$t, predict(fit), type='l', col='red')

## Transform to beta's
coef_transform = function(coef, index_range) N %*% coef[index_range]

t_int <- coef_transform(coef, seq(1, length=df))
tb_1 <- coef_transform(coef, seq(df + 1, length=df))
tb_2 <- coef_transform(coef, seq(2*df + 1, length=df))
y_hat <- t_int + (tb_1 * vcdata$x.1) + (tb_2 * vcdata$x.2)

coef_matrix <- data.frame(Time= vcdata$t, Int = t_int, B1 = tb_1, B2 =tb_2)
tidy_coef <- melt(coef_matrix, id="Time")

tidy_coef %>% ggplot(aes(x = Time, y = value, colour = variable)) +
  geom_line() + 
  ggtitle("Coefficient Values vs Time") + xlab("Time") + 
  ylab("Coefficient Vals") +
  theme_bw()
```


```{r}

cov <- vcov(fit)
covTransform <- function(index_range) {
  sqrt(diag(N %*% cov[index_range,index_range] %*% t(N)))
}

intCov <- covTransform(seq(1, length=df))
b1Cov <- covTransform(seq(df + 1, length=df))
b2Cov <- covTransform(seq(2*df + 1, length=df))

combineDfs <- function(coef, se) {
  cbind(time=coef_matrix$Time, 
        var=rep(coef, 500), 
        value=unname(coef_matrix[coef]),
        se=se)  
}

ptErrors <- as.data.frame(
  rbind(
  combineDfs('Int', intCov),
  combineDfs('B1', b1Cov),
  combineDfs('B2', b2Cov)
  ))

ptErrors %>% ggplot(aes(x = time, y = value)) +
  geom_line(aes(color = var, group = var)) +
  geom_errorbar(aes(ymin = value - 2*se, ymax = value + 2*se, color = var), 
                alpha=0.3) +
  ggtitle("Coefficient Values with Pointwise SEs") + xlab("Time") + 
  ylab("Coefficient Vals") +
  theme_bw()

```

